Carbonate chemistry at Ny-Ålesund

### 2021-12-29 awipev-CO2.Rmd
cleaning up continued

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", xlab = "x", 
                   ylab = "y", proplot = TRUE, ...) 
  { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
predicted <- fit$regression.results$Intercept[1] + xdata * fit$regression.results$Slope[1]
error <- xdata - predicted
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = title,
       subtitle = paste("\ntype: ", type,
                     "; R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "\nP =", signif(fit$P.param, 3),
                    "rmse =", signif(sqrt(mean(error^2, na.rm=TRUE)), 3)),
       x = xlab,
       y = ylab) +
    theme_bw() +
    theme(plot.title = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8))
    if (proplot == TRUE) { # property-property plot
      plt <- plt +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        theme(aspect.ratio = 1)
    }
  return(list(plot = plt, fit = fit))
  } else {
  return(fit)
    }
}

@@

d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  # data post 2020-03-01 eliminated because no maintenance and lots of algae
  dplyr::mutate(pco2_fb = ifelse(datetime >= as.POSIXct("2020-03-01 00:00"), NA, pco2_fb)) %>%
  arrange(datetime)

@@

```{r durafet calibration, echo=FALSE, message= FALSE, warning=FALSE}
# deployment period between 2 spectro measurements.
durPeriod <- d_all%>%
  dplyr::select(datetime, ph_s_dur_temp_fb)%>%
  dplyr::filter(!is.na(ph_s_dur_temp_fb))

durPeriod <-  durPeriod %>%
  dplyr::arrange(datetime) %>%
  dplyr::mutate(middle=  seconds_to_period(c(0,difftime(time1= datetime[-1], time2= datetime[-nrow(durPeriod)])/2* 3600 )))%>%
  dplyr::mutate(dur_interval = datetime + middle[2:nrow(durPeriod)])

@@

# Offset between durafet pH and spec pH 
duroffset <- left_join(durPeriod%>%dplyr::select(datetime,ph_s_dur_temp_fb), d_all%>%dplyr::select(datetime,ph_dur ), by="datetime")
dat_dur <- mutate(duroffset, diff_pH_dur = ph_s_dur_temp_fb - ph_dur)
average_offset_dur <- round(mean(dat_dur$diff_pH_dur, na.rm=TRUE), digits = 2)
sd_offset_dur <- round(sd(abs(dat_dur$diff_pH_dur), na.rm=TRUE), digits = 2)
n_offset_dur <- length(!is.na(dat_dur$diff_pH_dur))

plot_offset_dur <- ggplot(data = filter(dat_dur, !is.na(diff_pH_dur))) +
  geom_hline(yintercept=mean(dat_dur$diff_pH_dur, na.rm = TRUE), size=0.3, col="red") +
  annotate("text", x = dat_dur$datetime[3], y =-0.16, label = "offset mean", color="red")+
  geom_hline(yintercept=0, size=0.3) +
  geom_point(aes(x = datetime, y = diff_pH_dur), size = 1, colour = "#51C56AFF") +
  labs(x = NULL, y=expression(paste(pH[T], " difference")), title = NULL) +
  scale_x_datetime(date_breaks="month", date_minor_breaks="1 week", labels=date_format("%b %Y")) +
  Mytheme() + theme(panel.grid.major.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))

@@

# RANGE of pH
range_dur <- tbl_df(
  d_all %>%
  group_by(date, montht, year=year(datetime)) %>%
  summarize(pH_range = max(ph_dur_corr) - min(ph_dur_corr))
)
quant <- quantile(range_dur$pH_range, probs =c(2.5/100, 97.5/100), na.rm =TRUE)

# Per month
range_month <-
  range_dur %>%
  group_by(montht) %>%
  summarize(pH_range_month = mean(pH_range, na.rm=TRUE))

size_point <- 0.4 # size of data points
p <-
  ggplot(data = tidyr::drop_na(range_dur), 
         aes(x = as.factor(montht), y = pH_range)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(y =expression(paste(pH[T], " range durafet")), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
print(p) # plot_range_dur
rm(durafet_contros_xts)
```

@@

```{r Range per month, echo=FALSE, message= FALSE, warning=FALSE}
my_skim <- skim_with(numeric = sfl(ave = ~ mean(., na.rm = TRUE),
                                   med = ~ median(., na.rm = TRUE),
                                   min = ~ min(., na.rm = TRUE),
                                   max = ~ max(., na.rm = TRUE),
                                   q5 = ~ quantile(., 0.05, na.rm = TRUE),
                                   q95 = ~ quantile(., 0.95, na.rm = TRUE)
                                   ), 
                     append = FALSE)

@@

### **seaFET / durafet pHs (all > 8 m)**

```{r pHs comparaison 4°C, echo=FALSE, message= FALSE, warning=FALSE}
# prepare data and remove sf biofouling period
# TODO consider biofouling period elsewhere too
#   pH_sf <- d_all %>%
#   dplyr::select(datetime, sal_insitu_ctd_10m, temp_insitu_ctd_10m, pHint_tot_sf_temp_4, depth_group) %>%
#   #tidyr::drop_na() %>%
#   dplyr::filter(datetime < as.POSIXct("2018-01-01") | datetime > as.POSIXct("2018-02-03")) %>% 
#   dplyr::mutate(pHint_tot_sf_temp_4_10m = ifelse(depth_group == ">8", pHint_tot_sf_temp_4, NA))
# 
# #pH_sf$pHint_tot_sf_temp_4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_10m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_10m, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
# 
# pH_dur <- d_all %>%
#   dplyr::select(datetime, sal_fb,temp_fb, pH_dur)%>%
#   tidyr::drop_na() %>%
# pH_dur$pH_dur_temp_4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
# 
# # add pH sf and pH dur normalized at 4°C ti d_all
# #d_all <- left_join(d_all, pH_sf%>%dplyr::select(datetime, pHint_tot_sf_temp_4 ), by="datetime")
# d_all <- d_all %>% 
#   left_join(pH_dur %> %dplyr::select(datetime, pH_dur_temp_4), by="datetime")
lm <- lm2(x = d_all$pH_dur_temp_4, y = d_all$pHint_tot_sf_temp_4)
p <- d_all %>% 
  dplyr::select(pH_dur_temp_4, pHint_tot_sf_temp_4, monthd) %>% 
  drop_na() %>% 
  ggplot(aes(x = pH_dur_temp_4, y = pHint_tot_sf_temp_4))+
  geom_point(aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$intercept_MA, slope = lm$slope_MA), colour = "blue") +
  coord_fixed(ratio = 1 , xlim=c(7.8, 8.35), ylim=c(7.8, 8.35)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$rsquare, 3),
                     "; Intercept =", signif(lm$intercept_MA, 3),
                     "; Slope =", signif(lm$slope_MA, 3),
                     "; P =", signif(lm$P, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  labs( x = "pH durafet, 4°C", y = "pH seaFET, 4°C", color = "Months") + theme_bw()
print(p)

@@

### **Calculated / Durafet pHs**

```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize ph_calc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb_mix, temp_fb, ph_calc_fb, at)%>%
  dplyr::filter(!is.na(ph_calc_fb), !is.na(temp_fb), !is.na(sal_fb_mix))
pH_calc$pH_calc_fb_temp_4 <- pHinsi(pH = pH_calc$ph_calc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb_mix, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

# join pH calc norm + pH dur/sf norm in d_all
d_all <- left_join(d_all, pH_calc %>% dplyr::select(datetime, pH_calc_fb_temp_4), by="datetime")

@@

### **Calculated / seaFET pHs**

```{r pHs comparaison sf calc, echo=FALSE, message= FALSE, warning=FALSE}
seafet_calc_contros_xts <- dplyr::select(d_all, datetime, pHint_tot_sf_temp_4, pH_calc_fb_temp_4)
seafet_calc_contros_xts <- as.xts(seafet_calc_contros_xts, order.by = d_all$datetime)
dygraph(seafet_calc_contros_xts, group = "awipev", main="pH calculated vs pH seaFET, normalized at 4°C", ylab="pHT") %>%
    dySeries("pHint_tot_sf_temp_4", label = "pH seaFET", color ="blue", strokeWidth=0, pointSize=2) %>%
    dySeries("pH_calc_fb_temp_4", label = "pH calculated", color =  "red", strokeWidth = 0, pointSize = 4 ) %>%
    # dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
    dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
       dyLegend(show = "follow")   %>%
dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
    dyRangeSelector(height = 30, dateWindow= NULL)
rm(seafet_calc_contros_xts)
```

### **Spectro / durafet**

```{r pHs comparaison durafet spectro, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / durafet
lm <- lm2(x = d_all$ph_s_dur_temp_fb, y = d_all$pH_dur)
p <- d_all %>% 
  dplyr::select(ph_s_dur_temp_fb, pH_dur, monthd) %>% 
  drop_na() %>% 
  ggplot(aes(x = ph_s_dur_temp_fb, y= pH_dur, label = monthd)) +
  geom_text( aes(color=factor(monthd))) + 
  scale_color_discrete(guide = "none")+
  geom_abline(slope=1, intercept =0, linetype = "dashed") +
  geom_abline(aes(intercept = lm$intercept_MA, slope = lm$slope_MA),
              colour = "blue") +
  labs(title = paste("pH spectro vs pH durafet (both at FB temperature","\n Adj R2 = ", signif(lm$rsquare, 3),
                     "; Intercept =", signif(lm$intercept_MA, 3),
                     "; Slope =", signif(lm$slope_MA, 3),
                     "; P =", signif(lm$P, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme_bw() +
  theme(aspect.ratio = 1, plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 , xlim=c(8,8.4), ylim=c(8,8.4))+
  labs(x="pH durafet", y="pH spectro", color = "Months")
print(p)
```

>>>>

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", xlab = "x", 
                   ylab = "y", proplot = TRUE, ...) 
  { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
predicted <- fit$regression.results$Intercept[1] + xdata * fit$regression.results$Slope[1]
error <- xdata - predicted
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = title,
       subtitle = paste("\ntype: ", type,
                     "; R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "\nP =", signif(fit$P.param, 3),
                    "rmse =", signif(sqrt(mean(error^2, na.rm=TRUE)), 3)),
       x = xlab,
       y = ylab) +
    theme_bw() +
    theme(plot.title = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8))
    if (proplot == TRUE) { # property-property plot
      plt <- plt +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
        theme(aspect.ratio = 1) +
        coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on")
    }
  return(list(plot = plt, fit = fit))
  } else {
  return(fit)
    }
}

@@

d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  # pCO2 data post 2020-03-01 eliminated because no maintenance and lots of algae
  # pH durafet data post 2020-01-01 eliminated because no maintenance and lots of algae
  dplyr::mutate(pco2_fb = ifelse(datetime >= as.POSIXct("2020-03-01 00:00"), NA, pco2_fb),
                ph_dur = ifelse(datetime >= as.POSIXct("2020-01-01 00:00"), NA, ph_dur)) %>%
  arrange(datetime)

@@

```{r durafet calibration, echo=FALSE, message= FALSE, warning=FALSE}
# deployment period between 2 spectro measurements.
# TODO the following throws a warning but seems to work
durPeriod <- d_all %>%
  dplyr::select(datetime, ph_s_dur_temp_fb) %>%
  dplyr::filter(!is.na(ph_s_dur_temp_fb))
durPeriod <- durPeriod %>% 
  #dplyr::arrange(datetime) %>%
  dplyr::mutate(middle =  seconds_to_period(
   c(0, difftime(time1 = datetime[-1], time2 = datetime[-nrow(durPeriod)])/2* 3600 ))) %>%
  dplyr::mutate(dur_interval = datetime + middle[2:nrow(durPeriod)])

@@

# Offset between durafet pH and spec pH 
duroffset <- left_join(durPeriod %>% dplyr::select(datetime, ph_s_dur_temp_fb), d_all %>% dplyr::select(datetime, ph_dur), by="datetime")
dat_dur <- duroffset %>% 
  drop_na() %>% 
  dplyr::mutate(diff_pH_dur = ph_s_dur_temp_fb - ph_dur)
average_offset_dur <- round(mean(dat_dur$diff_pH_dur, na.rm=TRUE), digits = 2)
sd_offset_dur <- round(sd(abs(dat_dur$diff_pH_dur), na.rm=TRUE), digits = 2)
n_offset_dur <- length(!is.na(dat_dur$diff_pH_dur))

plot_offset_dur <- dat_dur %>% 
  drop_na() %>% 
  ggplot() +
  geom_hline(yintercept = mean(dat_dur$diff_pH_dur, na.rm = TRUE), size=0.3, col="red") +
  annotate("text", x = dat_dur$datetime[3], y =-0.16, label = "offset mean", color="red")+
  geom_hline(yintercept=0, size=0.3) +
  geom_point(aes(x = datetime, y = diff_pH_dur), size = 1, colour = "#51C56AFF") +
  labs(x = NULL, y=expression(paste(pH[T], " difference")), title = NULL) +
  scale_x_datetime(date_breaks="month", date_minor_breaks="1 week", labels=date_format("%b %Y")) +
  Mytheme() + theme(panel.grid.major.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Range of pH per day
range_dur <- d_all %>%
  group_by(date, montht) %>%
  summarize(pH_range = max(ph_dur_corr) - min(ph_dur_corr))
quant <- quantile(range_dur$pH_range, probs =c(2.5/100, 97.5/100), na.rm =TRUE)
# Range of pH per month
range_month <- range_dur %>%
  group_by(montht) %>%
  summarize(pH_range_month = mean(pH_range, na.rm=TRUE))

p <- range_dur %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x = as.factor(montht), y = pH_range)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(y =expression(paste(pH[T], " range durafet")), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
print(p) # plot_range_dur
rm(durafet_contros_xts)
```

@@

```{r Range per month, eval = FALSE, echo=FALSE, message= FALSE, warning=FALSE}
my_skim <- skim_with(numeric = sfl(ave = ~ mean(., na.rm = TRUE),
                                   med = ~ median(., na.rm = TRUE),
                                   min = ~ min(., na.rm = TRUE),
                                   max = ~ max(., na.rm = TRUE),
                                   q5 = ~ quantile(., 0.05, na.rm = TRUE),
                                   q95 = ~ quantile(., 0.95, na.rm = TRUE)
                                   ), 
                     append = FALSE)

@@

### **seaFET vs durafet pHs at 4°C (all > 8 m)**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
reg <- ggreg(
  xdata = d_all$pH_dur_temp_4,
  ydata = d_all$pHint_tot_sf_temp_4,
  type = "MA",
  plot = TRUE,
  title = "seaFET vs durafet pH at 4 °C",
  xlab = "pH durafet, 4 °C",
  ylab = "pH seaFET, 4 °C",
  proplot = TRUE
)
print(reg$plot + xlim(c(7.8, 8.35)) + ylim(c(7.8, 8.35)))

@@

### **Calculated vq Durafet**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize pH_calc from AT/CT to 4°c
tmp <- discrete %>%
  dplyr::select(datetime, sal_fb_mix, temp_fb, ph_calc_fb, at) %>%
  tidyr::drop_na()
tmp$pH_calc_fb_temp_4 <- pHinsi(pH = tmp$ph_calc_fb, ALK=tmp$at*1e-6, Tinsi=4, Tlab=tmp$temp_fb, Pinsi=11/10, S=tmp$sal_fb_mix, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

# join pH calc norm + pH dur/sf norm in d_all
d_all <- left_join(d_all, tmp %>% dplyr::select(datetime, pH_calc_fb_temp_4), by="datetime")

@@

### **Calculated vs seaFET**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
seafet_calc_contros_xts <- dplyr::select(d_all, datetime, pHint_tot_sf_temp_4, pH_calc_fb_temp_4)
seafet_calc_contros_xts <- as.xts(seafet_calc_contros_xts, order.by = d_all$datetime)
dygraph(seafet_calc_contros_xts, group = "awipev", main="pH calculated vs pH seaFET, normalized at 4°C", ylab="pHT") %>%
    dySeries("pHint_tot_sf_temp_4", label = "pH seaFET", color ="blue", strokeWidth=0, pointSize=2) %>%
    dySeries("pH_calc_fb_temp_4", label = "pH calculated", color =  "red", strokeWidth = 0, pointSize = 4 ) %>%
    # dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
    dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
       dyLegend(show = "follow")   %>%
dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
    dyRangeSelector(height = 30, dateWindow= NULL)
rm(seafet_calc_contros_xts)
```

### **Spectro vs durafet**
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / durafet
reg <- ggreg(
  xdata = d_all$ph_s_dur_temp_fb,
  ydata = d_all$pH_dur,
  type = "MA",
  plot = TRUE,
  title = "pH spectro vs pH durafet (both at FB temperature)",
  xlab = "pH durafet",
  ylab = "pH spectro",
  proplot = TRUE
)
print(reg$plot + xlim(c(7.8, 8.35)) + ylim(c(7.8, 8.35)))
```
